---
title: "R Notebook"
output: html_notebook
---

```{r}
# Set the seed for reproducibility
#set.seed(123)

# Set the sample size
n <- 400

# Set the effect size (Cohen's d) for the three-way interaction
interaction_effect_size <- 0.10

# Generate data for the three continuous variables
x1 <- rnorm(n, mean = 0, sd = 1)
x2 <- rnorm(n, mean = 0, sd = 1)


num_simsPer <- 1000
alpha <- 0.05


powerMat <- vector(length=num_sims)

for(j in 1:num_sims){
  
  # Initialize a counter for the number of times the null hypothesis is rejected
  rejections <- 0

  # Run the simulations
  for (i in 1:num_sims) {
    
    # Generate the response variable with the three-way interaction
    y <- 1 + 1*x1 + 1*x2 + (x1*x2)*interaction_effect_size + rnorm(n, mean = 0, sd = 2)
  
    # Create a dataframe with the variables
    data <- data.frame(y, x1, x2, x3)
    
    # Fit the linear regression model on the training set
    model <- lm(y ~ x1*x2, data = data)
    
    # Perform the hypothesis test for the three-way interaction term
    p_value <- summary(model)$coefficients[4,4]
    
    # Check if the p-value is less than the significance level
    if (p_value < alpha) {
      rejections <- rejections + 1
    }
    
  }

# Calculate the power of the test
power <- rejections / num_sims

powerMat[j] <- power

}
```

```{r}
# Set the seed for reproducibility
#set.seed(123)

# Set the sample size
n <- 400

# Set the effect size (Cohen's d) for the three-way interaction
interaction_effect_size <- 0.10

# Generate data for the three continuous variables
x1 <- rnorm(n, mean = 0, sd = 1)
x2 <- rnorm(n, mean = 0, sd = 1)
x3 <- rnorm(n, mean = 0, sd = 1)

# Set the number of simulations and the false positive rate
num_sims <- 1000
alpha <- 0.05

# Initialize a counter for the number of times the null hypothesis is rejected
rejections <- 0

# Run the simulations
for (i in 1:num_sims) {
  
  # Generate the response variable with the three-way interaction
  y <- 1 + 1*x1 + 1*x2 + 1*x3 + (x1*x2*x3)*interaction_effect_size + rnorm(n, mean = 0, sd = 2)

  # Create a dataframe with the variables
  data <- data.frame(y, x1, x2, x3)
  
  # Fit the linear regression model on the training set
  model <- lm(y ~ x1 + x2 + x3 + x1*x2 + x1*x3 + x2*x3 + x1*x2*x3, data = data)
  
  # Perform the hypothesis test for the three-way interaction term
  p_value <- summary(model)$coefficients[8,4]
  
  # Check if the p-value is less than the significance level
  if (p_value < alpha) {
    rejections <- rejections + 1
  }
  
}

# Calculate the power of the test
power <- rejections / num_sims

# Print the power
print(power)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

